---
title: "ggmap"
author: "Spencer Mathews"
date: "`r Sys.Date()`"
output: html_document
---


- ggmap / [CRAN](CRAN - Package ggmap) / [GitHub](https://github.com/dkahle/ggmap)
- (ggmap: Spatial Visualization with
ggplot2 / by David Kahle and Hadley Wickham)[https://journal.r-project.org/archive/2013-1/kahle-wickham.pdf]

### Tutorials

- [Geographic visualization with R's ggmap](http://blog.dominodatalab.com/geographic-visualization-with-rs-ggmaps/)

---

The workhorse function is `get_map`, which is a wrapper for `get_googlemap`, `get_openstreetmap`, `get_stamenmap`, and `get_cloudmademap`. It downloads images and formats them so `ggmap` can plot them (using `ggplot2`). `qmap` combines these steps.

Arguments include `location`, `zoom`
- `location` - longitude/latitude or string which is passed to `geocode`
(from 3 to 20).
- `zoom`
- `source`/`maptype`

> location an address, longitude/latitude pair (in that order), or left/bottom/right/top bounding
box
zoom map zoom, an integer from 3 (continent) to 21 (building), default value 10 (city).
openstreetmaps limits a zoom of 18, and the limit on stamen maps depends
on the maptype. "auto" automatically determines the zoom for bounding box
specifications, and is defaulted to 10 with center/zoom specifications. maps of
the whole world currently not supported.
scale scale argument of get_googlemap or get_openstreetmap
maptype character string providing map theme. options available are "terrain", "terrainbackground",
"satellite", "roadmap", and "hybrid" (google maps), "terrain", "watercolor",
and "toner" (stamen maps), or a positive integer for cloudmade maps
(see ?get_cloudmademap)
source Google Maps ("google"), OpenStreetMap ("osm"), Stamen Maps ("stamen"), or
CloudMade maps ("cloudmade")
force force new map (don’t use archived version)
messaging turn messaging on/off
16 get_navermap
urlonly return url only
filename destination file for download (file extension added according to format)
crop (stamen and cloudmade maps) crop tiles to bounding box
color color ("color") or black-and-white ("bw")
language language for google maps
api_key an api key for cloudmade maps



ggmap

> ggmap Plot a ggmap object
Description
> ggmap plots the raster object produced by get_map.
Usage
ggmap(ggmap, extent = "panel", base_layer, maprange = FALSE,
legend = "right", padding = 0.02, darken = c(0, "black"), ...)
Arguments
ggmap an object of class ggmap (from function get_map)
extent how much of the plot should the map take up? "normal", "device", or "panel"
(default)
base_layer a ggplot(aes(...), ...) call; see examples
maprange logical for use with base_layer; should the map define the x and y limits?
legend "left", "right" (default), "bottom", "top", "bottomleft", "bottomright", "topleft",
"topright", "none" (used with extent = "device")
padding distance from legend to corner of the plot (used with legend, formerly b)
darken vector of the form c(number, color), where number is in [0, 1] and color is a
character string indicating

```{r, message=FALSE}
library(ggmap)
```

```{r}
qmap("Davis,CA", zoom = 14, source = "osm")
```


```{r, message=FALSE, cache=TRUE}
us <- c(left = -125, bottom = 25.75, right = -67, top = 49)
map <- get_map(us, zoom = 3, source="stamen", maptype = "toner-lite")
ggmap(map)
```



```{r}
map <- get_map("Atlanta", zoom=10)
p <- ggmap(map)
ggsave(p, file = "map7.png", width = 5, height = 5, type = "cairo-png")  
```

