---
title: "Learn ggplot2"
author: "Spencer Mathews"
date: "`r Sys.Date()`"
output: html_document
---

```{r, echo=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
```



You call qplot or ggplot and then add _layers_ with `+`.




Layers: can often use a `geom_*` or `stat_*` interchangeably.


# Quickstart

### Labeling 
`labs(title= , x= , y= )` or can be individually set with `ggtitle()`, `xlab()`, `ylab()`




## qplot

While ggplot requires a data frame qplot can be used to plot vector data.

```
qplot(x, y = NULL, ..., data, facets = NULL, margins = FALSE,
  geom = "auto", stat = list(NULL), position = list(NULL), 
  xlim = c(NA, NA), ylim = c(NA, NA), log = "", 
  main = NULL, xlab = deparse(substitute(x)), ylab = deparse(substitute(y)), asp = NA)
```

### Other Parameters to qplot




group

color
shape
size

`alpha=` give it a variable or a fixed value like `alpha=I(1/10)`

fill - learn about continuous vs discrete, and whether your things is a factor


## ggplot




## Saving Plots

`ggsave` - save plot




# Aesthetics (aes)
aes/aes_string

You might include this in the call the ggplot or as a layer.




---

# Geometries (geom)


### [`geom_line`](http://docs.ggplot2.org/current/geom_line.html)

### `geom_path` - preserves data ordering not numerical

### [`geom_smooth`](http://docs.ggplot2.org/current/geom_smooth.html)

By default it uses `stat_smooth`.


### `geom_ribbon`

Required aes are x, ymax, ymin.


### `geom_histogram`

> geom_histogram is an alias for `geom_bar` plus `stat_bin`

All of`geom_histogram()`, `stat_bin()`, and `geom_bar` default to `position='stack'`.


### `geom_bar`

Oddly defaults to `stat='bar'` and you get counts (don't set y).  If you want to represent values instead of counts set stat="identity" and supply y.

We want discrete/factor values here ideally.

For bar charts and histograms you can set `position='dodge'` to have different data series next to eachother vs stacked.


### `geom_density`

By default does not fill the plot, whereas `stat_density()` does.

Defaults to `position='identity'` but can also use `position='stack` or `position='fill'`, etc.


### `geom_area`

Continuous analog of stacked bar chart, and special case of geom_ribbon with a minimum of 0 and default position of stacked.


### `geom_errorbar`
Can explicitely set aes(ymin= , ymax= )


### `geom_pointrange`
### `geom_linerange`
### `geom_crossbar`


### `geom_tile`
### `geom_raster`

---

# Stats

### `stat_smooth`

method=lm/glm/loess/...
Possibly more options avialable than geom_smooth.
Defaults to show standard error but can also turn this off with se=FALSe
`formula` option lets you specify smoothing formula e.g. y ~ x, y ~ poly(x, 2), y ~ log(x)


### `stat_identity`


### `stat_summary`


Can specify summary functions in fun.data, fun.y, fun.ymax, fun.ymin from the Hmisc package:
mean_se - mean and standard errors
mean_cl_boot(x, ...) - confidence limits with nonparametric bootstrapping method, does not assume normality.
mean_cl_normal(x, ...) - sample mean and upper and lower Gaussian confidence limits based on t-distribution.
mean_sdl(x, ...) - mean plus or minus constant time standard deviation, mult=2 is default
median_hilow(x, ...) - sample median and some outlier quantiles



### [`stat_smooth`](http://docs.ggplot2.org/current/stat_smooth.html)

method=lm, glm, gam, loess, rlm


### `stat_bin`

The default statistic for `geom_histogram` and `geom_bar`.

Defaults to `geom="bar"` and `position="stack"`

Can adjust `binwidth=`.

To create unit area histogram use `stat_bin(aes(y = ..density..), geom="line", position="identity"))`.  The geom and position may not be necessary.  Can also use this `..density..` when calling geom_histogram.

Arguments:
`binwidth=`

mapping/aes


Returns special variables `..count..` and `..density..`.  Create a unit historgram by setting `y=..density..` so get a density instead of count.


### `stat_density()`

Defaults to `geom = "area"`, and `position = "stack"`
  
Can also set `adjust=` where higher values are smoother (default 1)

### [`stat_sum`](http://docs.ggplot2.org/current/stat_sum.html)

Cool summing plot.  Grouping specifies how dots will be scaled.

Can set `stat_sum(aes(group = 1))` to size based on proportions of groupings, and if desired `scale_size(range = c(3, 10))`.  Specifying another grouping will give proportions relative to that group.

Colour can be specified, too but it must be in addition to grouping or else the summing won't happen.
You can `weight` by another variable.
Show actual sums instead of proportions by `size = ..n..`, which might imply grouping?


# Scales

`scale_alpha()` = `scale_alpha_continuous`
`scale_alpha_discrete()`


# Position


### `position_identity`

### `position_dodge`

### `position_stack`

### `position_fill`

### `position_jitter`



# Coordinate Systems

### `coord_cartesian`

### `coord_polar`

### `coord_map`



---

# Facets

### `facet_wrap`

### `facet_grid`


---

# Scales

`guides`/`guide_colourbar`/`guide_legend`





---

# [Legends](http://www.cookbook-r.com/Graphs/Legends_%28ggplot2%29/) ([docs](http://docs.ggplot2.org/current/labs.html))


Turning off legends can be done with `guides(<color|fill|linetype|shape|etc>=FALSE)` and some other ways.  This may do other things too.

1. `scale_colour/fill_hue/discrete`
2. `scale_colour/fill_gradient/continuous`
3. `scale_shape[_continuous/discrete]`
4. `scale_size[_continous/discrete]`
3. etc.


---

`theme` - for lots of stuff affecting looks


---

# Gotchas
If outputting to file must use print(), eg. print(qlot(...))

`coord_cartesian(xlim=, ylim=)` - does not drop points
xlim/ylim - convenience functions, drops points



---

# Tips and Tricks


Check in about setting alpha using `I(1/2)` and see if the `I` is necessary and when.

### Means and Standard Deviations

The `stat_summary` layer can be used to plot an envelope around the mean.  It normally gives the standard error but you can specify `fun.data=mean_sdl`.  Using `geom='smooth'` plots the mean and fills a transparent ribbon around it.  It's not entirely clear how if any smoothing is actually done here.  Using `geom='ribbon'` just plots the surrounding ribbon and but you should specify and `alpha` if you want to add the mean line.  If using ribbon `alpha=4/10` seems to give the same transparency as smooth.


1. smooth
```
ggplot(a, aes(x=x,y=y)) + geom_point() +
  stat_summary(fun.data = 'mean_sdl', mult = 1, geom = 'smooth')
```

2. ribbon
```
ggplot(a, aes(x=x,y=y)) + 
  stat_summary(fun.y = 'mean', colour = 'blue', geom = 'line') +
  stat_summary(fun.data = 'mean_sdl', geom = 'ribbon', alpha = 0.2)
```



```
# Hadley's wrapper function
stat_sum_df <- function(fun, geom="crossbar", ...) { 
  stat_summary(fun.data=fun, colour="blue", geom=geom, width=0.2, ...) 
  
p + stat_sum_df("mean_cl_normal", geom = "smooth") 
} 
```

```
nmmaps$mino3<-nmmaps$o3run-sd(nmmaps$o3run, na.rm=T)
nmmaps$maxo3<-nmmaps$o3run+sd(nmmaps$o3run, na.rm=T)

ggplot(nmmaps, aes(date, o3run))+geom_ribbon(aes(ymin=mino3, ymax=maxo3), fill="steelblue2", color="steelblue2")+
  geom_line(color="steelblue4", lwd=1)
```






# Examples and Use Cases

## Histograms and Density

Histograms may be constructed with the `geom_histogram()` and `stat_bin()` layers, and density plots with `geom_density()` and `stat_density()`.

Generate data from two offset distributions to test variations.

```{r}
samples = data.frame(a=rnorm(1000, mean=0), b=rnorm(1000, mean=1))
melt_samples = samples %>% gather(grouping, value)
```


Both `geom_histogram()` and `stat_bin()` default to `position='stack'`.

```{r}
melt_samples %>% ggplot(aes(x=value)) + geom_histogram(aes(fill=grouping))
```

```{r}
melt_samples %>% ggplot(aes(x=value)) + stat_bin(aes(fill=grouping))
```

```{r}
melt_samples %>% ggplot(aes(x=value)) + geom_density(aes(color=grouping))
```

```{r}
melt_samples %>% ggplot(aes(x=value)) + stat_density(aes(fill=grouping))
```

```{r}
melt_samples %>% ggplot(aes(x=value)) + stat_density(aes(fill=grouping), position='fill')
```


