---
title: "Learn Markdown"
author: "Spencer Mathews"
date: "`r Sys.Date()`"
output: html_document
---

This is oriented toward authoring markdown in RStudio/knitr.

---


# Links

- [Rmarkdown Homepage](http://rmarkdown.rstudio.com)


- [Fast-track publishing using knitr](http://gforge.se/2013/12/fast-track-publishing-using-knitr-part-i/)
- [Fast-track publishing using the new R markdown](http://gforge.se/2014/07/fast-track-publishing-using-rmarkdown/)



# Relevant Libraries

- `rmarkdown`
- `knitr`
- `shiny`

- [htmlwidgets](http://www.htmlwidgets.org/) - javascript bindings for console, R Markdown, Shiny, etc.


---

# R Markdown

It's super easy to compile Rmd documents with R Studio, but 


### Code

Inline with backticks.  If you start the code segment with r space then it will execute (three backticks appears to be usable inline? but don't do it)
Blocks with three backticks or indent 4 spaces or one tab.  Putting `{r}` after the three backticks makes the code executable.

At least in RStudio the paragraphs are placed in `<p></p>` tags, and default rendering of inline code is with `<code><\code>` and render with grey background.  Executed inline code does not seem to be output, only the result.  Code blocks are wrapped in `<pre><code><\code><\code>` tags and have white backgrounds, and executable code chunks are rendered `<pre class="r"><code><\code><\code>` and have grey backgrounds.

Here are some knitr examples:

```
example of simple code block (without specifying {r}), it is wrapped in pre and code without a class specifier
```

```{r}
# with {r}, evaluated code block is grey, and is wrapped in pre and code tags with class="r" specified in pre
# the output it generates is white and is also wrapped in pre and code tags but does not have class="r"
getwd()
```

```{r, collapse=TRUE}
# with {r, collapse=TRUE}, evaluated code block is grey, and is wrapped in pre and code tags with class="r" specified in pre
# the output it generates is just appended to this
getwd()
```

```{r, eval=FALSE}
# with {r, eval=FALSE}, grey background and same tags as if evaluated
getwd()
```

Here are some examples of including the `pre` and `code` HTML directly:

Backticks can be included as `&#96;&#96;&#96;`  It seems that newlines are treated as spaces in `<code>` blocks that are not within `<pre>`.

Including this code:

<pre><code>&#60;pre class="r"&#62;&#60;code&#62;&amp#96;&amp#96;&amp#96;{r}
R CODE
&amp#96;&amp#96;&amp#96;
&#60;/code&#62;&#60;/pre&#62;
</code></pre>

Gives this output:

<pre class="r"><code>&#96;&#96;&#96;{r}
R CODE
&#96;&#96;&#96;
</code></pre>

Experiments with specifying HTML tags explicitely:

<pre>&#60;pre&#62;
# the pre tag without a class specifier has white background even without code tag
</pre>

<pre><code>&#60;pre&#62;&#60;code&#62;
# without a class specifier wrapping in pre and code tags yields a white background
</code></pre>

<pre class="r">&#60;pre class="r"&#62;
# the pre tag with class="r" but no code tag still has grey background
# the code tag seems redundant with pre, except to tell web crawlers that it's code
</pre>


<pre class="r"><code>&#60;pre class="r"&#62;&#60;code&#62;
# the rmarkdown rdocs include backticks in code blocks with explicite pre and code tags
# and with class="markdown", here class="r" but in both cases the output is grey
# but when class is specified and code tag used looks like you get syntax highlighting!
</code></pre>





<code>&#60;code&#62;
</code>



### Tables

`knitr::kable()`


---

## Output

### HTML

### PDF

I noticed that if you do not leave a blank line between text and a figure generating code block then the figure will be put on the same line and likely run off the page.  This affected PDF output, but not HTML.



----------------------------------------

# rmarkdown package

- [rmarkdown @github](https://github.com/rstudio/rmarkdown)

`rmarkdown::render()` will knit Rmd (using knitr::knit()) and produce output with pandoc.  You can render plain markdown (flavor support?) and it will skip the knitr.

Defaults to HTML.  You can specify one or more output formats in the YAML header, as arguments to render, or externally in `_output.yaml`.

Built in output formats:
- html_document
- pdf_document
- word_document
- md_document
- beamer_presentation
- ioslides_presentation
- slidy_presentation

There are different options available for different engines, including `toc: true`, `theme: xxx`, etc.





----------------------------------------



# knitr package

Both LyX and R Studio support knitr!

- [knitr](http://yihui.name/knitr/) [github](https://github.com/yihui/knitr)
- [knitr (CRAN)](http://cran.r-project.org/web/packages/knitr)
- [knitr in a nutshell](http://kbroman.org/knitr_knutshell)
- [Fast-track publishing using knitr](http://gforge.se/2013/12/fast-track-publishing-using-knitr-part-i)



## Comments

I think you can commend within `<!--` and `-->`, unclear if it's just for HTML or also PDF output.  When compiling for html surrounding an


## Options


### [Knitr Objects](http://yihui.name/knitr/objects)

- `opts_chunk` and `opts_current`: manages options for code chunks
- `opts_knit`: manages options for the knitr package
- `knit_hooks`: manages hook functions
- `knit_patterns`: manages regular expressions to extract R code from the input document
- `knit_engines`: functions to deal with other languages

---

### Global Options


<pre class="r"><code>&#96;&#96;&#96;{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
&#96;&#96;&#96;
</code></pre>

---

### [Chunk Options](http://yihui.name/knitr/options/)

<pre class="r"><code>&#96;&#96;&#96;{r [lablel], [options]}
CODE
&#96;&#96;&#96;
</code></pre>


labels?


#### Alternative Language Engines

Knitr supports many [language engines](http://yihui.name/knitr/demo/engines).

engine='python'
opts_chunk$set(engine = 'python')

Be aware that except for the default `engine='R'` all chunks are executed in separate sessions so sharing variables can be tricky.




#### Code Evaluation

- `eval=FALSE`
whether or not to evaluate chunk, can also selectively execute lines


#### Code Evaluation

- `echo=TRUE`

- `results='hide'` 
- `results=asis` can be used with `cat` to print backticks and other special characters

- `collapse`

- `warning=TRUE`/`error=TRUE`/`message=TRUE`
- `include=FALSE` whether to include output in document, still evaluated, only option that isn't cached



#### Code Decoration


#### Cache

- `cache=TRUE`

Do not cache chunks which load libraries needed by uncached chunks.
May need to explicitly set `cache=FALSE` on those chunks.


#### Plots


- `fig.path=` seems like will save the figures here
- `dev=` can output multiple formats like `=c('png','jpeg')` and all will go to fig path if set but looks like first is embedded.  If chunk cached might go there too. Options include `tikz` (which could be nice)...

- `fig.width=7/fig.height=5`

in inches (5x7 is default), though perhaps px or other can be specified.
and can also specify globally either in YAML header as:

```
---
...
output:
  html_document:
    fig_width: 6
    fig_height: 4
---
```

or as

<pre class="r"><code>&#96;&#96;&#96;{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width7, fig.height=5)
&#96;&#96;&#96;
</code></pre>



---

# Interactive Shiny

See notes on [Shiny](shiny.html).

