---
title: "Learn Markdown"
author: "Spencer Mathews"
date: "`r Sys.Date()`"
output: html_document
---


This is oriented toward authoring markdown in RStudio/knitr but will be more general than that.


# Flavors of Markdown



---

# Links

- [Rmarkdown Homepage](http://rmarkdown.rstudio.com)


# Relevant Libraries

- `rmarkdown`
- `knitr`
- `shiny`

- [htmlwidgets](http://www.htmlwidgets.org/) - javascript bindings for console, R Markdown, Shiny, etc.


# R Markdown




### Code
Inline with backticks.  If you start the code segment with r space then it will execute (three backticks appears to be usable inline? but don't do it)
Blocks with three backticks or indent 4 spaces or one tab.  Putting `{r}` after the three backticks makes the code executable.

At least in RStudio the paragraphs are placed in `<p></p>` tags, and default rendering of inline code is with `<code><\code>` and render with grey background.  Executed inline code does not seem to be output, only the result.  Code blocks are wrapped in `<pre><code><\code><\code>` tags and have white backgrounds, and executable code chunks are rendered `<pre class="r"><code><\code><\code>` and have grey backgrounds.

Here are some knitr examples:

```
example of simple code block, it is wrapped in pre and code without a class specifier
```

```{r}
# evaluated code block is grey, and is wrapped in pre and code tags with class="r" specified in pre
# the output it generates is white and is also wrapped in pre and code tags but does not have class="r"
getwd()
```


```{r, eval=FALSE}
# unevaluated R code block has grey background and while this is has class="r" the evaluated does not have a class specified
getwd()
```

Here are some examples of including the `pre` and `code` HTML directly:

Backticks can be included as `&#96;&#96;&#96;`  It seems that newlines are treated as spaces in `<code>` blocks that are not within `<pre>`.

Including this code:

<pre><code>&#60;pre class="r"&#62;&#60;code&#62;&amp#96;&amp#96;&amp#96;{r}
R CODE
&amp#96;&amp#96;&amp#96;
&#60;/code&#62;&#60;/pre&#62;
</code></pre>

Gives this output:

<pre class="r"><code>&#96;&#96;&#96;{r}
R CODE
&#96;&#96;&#96;
</code></pre>

Experiments:

<pre>&#60;pre&#62;
# the pre tag without a class specifier has white background even without code tag
</pre>

<pre><code>&#60;pre&#62;&#60;code&#62;
# without a class specifier wrapping in pre and code tags yields a white background
</code></pre>

<pre class="r">&#60;pre class="r"&#62;
# the pre tag with class="r" but no code tag still has grey background
# the code tag seems redundant with pre, except to tell web crawlers that it's code
</pre>


<pre class="r"><code>&#60;pre class="r"&#62;&#60;code&#62;
# the rmarkdown rdocs include backticks in code blocks with explicite pre and code tags
# and with class="markdown", here class="r" but in both cases the output is grey
# but when class is specified and code tag used looks like you get syntax highlighting!
</code></pre>





<code>&#60;code&#62;
</code>




---

## Output

### HTML

### PDF

I noticed that if you do not leave a blank line between text and a figure generating code block then the figure will be put on the same line and likely run off the page.  This affected PDF output, but not HTML.


----------------------------------------

# [knitr](http://yihui.name/knitr/)

- [knitr in a nutshell](http://kbroman.org/knitr_knutshell)


## Comments

I think you can commend within `<!--` and `-->`, unclear if it's just for HTML or also PDF output.  When compiling for html surrounding an

## Options

### [Knitr Objects](http://yihui.name/knitr/objects)

- `opts_chunk` and `opts_current`: manages options for code chunks
- `opts_knit`: manages options for the knitr package
- `knit_hooks`: manages hook functions
- `knit_patterns`: manages regular expressions to extract R code from the input document
- `knit_engines`: functions to deal with other languages


### Global Options

```
\```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
\```
```

### Chunks

chunk labels?

### Chunk Options

`{r [lablel], [options]}`


#### `fig.width=7/fig.height=5`

in inches (5x7 is default), though perhaps px or other can be specified.
an also specify globally either in YAML header as:

```
output:
  html_document:
    fig_width: 6
    fig_height: 4
```

or as

```
{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width7, fig.height=5)
```

#### `eval=FALSE`
whether or not to evaluate chunk, can also selectively execute lines

#### `echo=TRUE`

#### `results='hide'`

`results=asis` can be used with `cat` to print backticks and other special characters

#### `warning=TRUE`/`error=TRUE`/`message=TRUE`

#### `cache=TRUE`
Do not cache chunks which load libraries needed by uncached chunks.
May need to explicitly set `cache=FALSE` on those chunks.

#### `include=FALSE`



---

# Interactive Shiny


